Catchy Title Goes Here

### 2018-06-20 Paper/paper.tex
updates to discussion text

The age-specific profile of immunity to measles infection can be a powerful tool for the assessment of vaccination programs and the planning for vaccine-based interventions [REF - Winter; Cutts and Hanson]. Comprehensive serological surveys can be prohibitively expensive and are thus only conducted infrequently.  Age-specific measles case incidence, by contrast, is routinely collected in most national health systems and has been previously used to infer age-specific infection risk via the catalytic model [REF - Griffiths; Grenfell and Anderson; Whittaker and Farrington; Ferrari 2010]. Here we present a novel extension of the catalytic model that uses age-specific vaccination coverage surveys to account for the competing age-specific rates of immunization via vaccination and natural infection.  We show, using a simulation experiment, that this model can be used to estimate age-specific seroprevalence using age-specific case incidence and vaccination coverage data. A necessary component of this model is the vaccine effectiveness, which gives the proportion of individuals who report vaccination on a survey that are effectively immunized by vaccination.  We show that, while this vaccine effectiveness is difficult to estimate in absolute value, our model-based estimator is positively correalted with the true value used in simulation; thus this estimate may be useful as a relative measure of vaccine effectiveness.  We further show that the inclusion of limited serological survey data, here a survey of children between 6-59 months of age, results in improvements of all model parameters, including vaccine effectiveness.

@@

The 2014 DHS in DRC included a sero survey that was limited to children between the ages of 6-59 months.

That survey provides only a limited perspective on the immunity in older age classes.  The case data, however DO provide this information.  This raises the question of what is the marginal benefit of conducting a sero survey.

very little marginal benefit in terms of assessing relative risk -- the age-specfic case data and vaccination coverage data do a good job of estimating the rank order of risk.

However, as illustrated in the simulation study, the sero survey allows more confident estimates of the rank order of vaccine effectiveness.

Limitations -- the model requires some significant assumptions that are almost certainly violated.

>>>>

The age-specific profile of immunity to measles infection can be a powerful tool for the assessment of vaccination programs and the planning for vaccine-based interventions [REF - Winter; Cutts and Hanson]. Comprehensive serological surveys can be prohibitively expensive and are thus only conducted infrequently.  Age-specific measles case incidence, by contrast, is routinely collected in most national health systems and has been previously used to infer age-specific infection risk via the catalytic model [REF - Griffiths; Grenfell and Anderson; Whittaker and Farrington; Ferrari 2010]. Here we present a novel extension of the catalytic model that uses age-specific vaccination coverage surveys to account for the competing age-specific rates of immunization via vaccination and natural infection.  We show, using a simulation experiment, that this model can be used to estimate age-specific seroprevalence using age-specific case incidence and vaccination coverage data. A necessary component of this model is the vaccination effectiveness, which gives the proportion of individuals who report vaccination on a survey that are effectively immunized by vaccination.  We show that, while this vaccination effectiveness is difficult to estimate in absolute value, our model-based estimator is positively correalted with the true value used in simulation; thus this estimate may be useful as a relative measure of vaccination effectiveness.  We further show that the inclusion of limited serological survey data, here a survey of children between 6-59 months of age, results in improvements of all model parameters, including vaccination effectiveness.

@@

The 2014 DHS in DRC included a sero-survey that was limited to children between the ages of 6-59 months. While this survey, by itself, can provide only a limited view of the total population immunity (including individuals older than 59 months) we have shown that combining these data with age-specific case incidence, over all ages, and vaccination coverage data can yield unbiased estimates of the age-specific seroprevalence across all ages, and thus, of the total susceptible population. In most provinces in DRC, these estimates did not differ greatly from those made without the inclusion of the sero-survey; however, in provinces where the estimates based on case incidence and vaccination coverage data only were biased, they tended to over-estimate population immunity.  Thus, even a relatively small investement in serological surveys may have benefits for population-level estimates.  Further, as illustrated in the simulation study, the inclusion of the serological data yields estimates of the vaccination effectiveness that are negatively biased, but preserve the rank order of estimates and thus may provide a useful relative measure of either vaccine program performance or confidence in vaccination coverage surveys.  While it is impossible to definitively identify whether the disparity between coverage and immunity is due to field effectiveness of the vaccine or weak recall in vaccination surveys, the resulting estimate could provide a measure of confidence in vaccine coverage values that may be of use in vaccine planning. The estimates of vaccination effectiveness for the provinces of DRC exhibit wide variation. Bandundu, Sud Kivu, Kasai Occidental, and Kasai Oriental rank as the 4 lowest estimates of vaccination effectiveness.  Again, it is impossible to determine whether this low estimate is due to low performance of the vaccination program or over-estimation of vaccination by monthers during the survey, these results suggest the value of further evaluation in these regions to evaluate vaccination levels.

This study has several limitations that should be considered when interpreting these results.  The first set of limitations pertain to the assumptions of the catalytic model. An implicit assumption of the catalytic model (and previous incarnations of the catalytic model [REF Griffiths, Grenfell and Anderson, Whittaker and Farrington] is that both the vaccination and FOI hazards have beeen fixed and constant in the past.  This assumption is almost certainly violated in practice as coverage with measles containing vaccine has been increasing in most countries in the world [REF Dabbagh et al, Progress to elimination WMMR]; thus, while the shape of the vaccine hazard may remain constant, the magnitude is likely to be increasing.  This implies that the fitted hazard is a reflection of the vaccination hazard over recent time, rather than current vaccination.  Similarly, we would expect the magnitude of FOI would be negatively correlated with vaccination coverage, and thus would be declining over time, rather than constant.  We note that in our simulations, vaccination coverage increased linearly over time from the time of introduction to the time of sampling and the fits to the simulated data still gave reasonable estimates of the age-specific seroprevalence at the time of sampling. In addition, the model, as we have stated it here does not formally account for the additional immunity in specific age cohorts that would result from pulsed, age-targeted vaccination campaigns (SIAs). We note that it is possible to include such pulsed campaigns in such an analysis per methods described in [Li et al PLoS Med].

A second set of assumptions arise from the inherent limitations of surveillance and survey data.
